Linear Ordered Attribute Grammar scheduling using SAT-solving

### 2014-10-10 paper/paper.tex
Updated abstract by Koen

\begin{abstract}
Attribute grammars are a formalism for describing tree computations (catamorphisms), and are therefore suitable for implementing the semantics of programming languages. A well-known subclass is the class of linear ordered attribute grammars, for which a linear evaluation order can be found statically. The problem of finding such an order is however known to be NP-hard.

In this paper we encode linear ordered attribute grammar scheduling as a SAT-problem. We encode cycle-freeness of the dependency graphs efficiently by making the graphs chordal, which is a new approach for encoding such graph property in SAT. We show that using Minisat as an external SAT-solver the scheduling algorithm runs much faster than existing scheduling algorithms on real-world examples. Furthermore, this approach allows fine-grained control over the resulting schedule, thereby enabling new attribute grammar optimizations.

>>>>

\begin{abstract}
Many computations over trees can be expressed using attribute
grammars. Compilers for attribute grammars need to find an evaluation
order (or schedule) in order to generate efficient code. For the class
of linear ordered attribute grammars such a schedule can be found
statically, but this problem is known to be NP-hard.

In this paper, we show how to encode linear ordered attribute grammar
scheduling as a SAT-problem. Cycle-freeness of the dependency graphs
is expressed by making the graphs chordal, which is a new approach for
encoding such graph property in SAT.

There are two main advantages to using a SAT-solver for scheduling:
(1) the scheduling algorithm runs much faster than existing scheduling
algorithms on real-world examples, and (2) by adding extra constraints
we obtain fine-grained control over the resulting schedule, thereby
enabling new scheduling optimizations.
